Detection of slow slip events using wavelet analysis of GNSS recordings

### 2022-06-29 paper/answer_reviewers.tex

\documentclass[letterpaper, 12pt]{article}

\usepackage{caption}
\usepackage{float}
\usepackage[T1]{fontenc}
\usepackage[lmargin=1 in, rmargin=1 in, tmargin=1 in, bmargin=1 in]{geometry}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{times}
\usepackage{xcolor}

\begin{document}

\textbf{Associate Editor}

\textit{Both reviewers expressed that the proposed MODWT method for slow-slip detection shows potential, but were concerned about whether the method could be applied successfully to regions without existing high-quality slow-slip or tremor catalogs.  An existing catalog appears to be necessary to select the detection threshold from the ROC diagram, and to identify false detections.  One way to address this concern would be, as Reviewer \#1 suggests, by applying the method to another subduction zone.  Alternatively, as Reviewer \#2 suggests, the manuscript could be revised to reframe the proposed method.  Successfully applying the method to another subduction zone would make the strongest case for the MODWT method, although it would be more work.  Either of these options for revision are acceptable to me, so I leave the choice to the authors.}

@@

\textit{This manuscript provide a useful and much needed exploration and explanation of the relationship between wavelet levels and offset duration. In particular, advocating for the use of the MODWT over the standard discrete wavelet transform provides a nice way to avoid boundary effects at the edge of each wavelet octave and provides better temporal location for identification of wavelet peaks. The applicability to the case of Cascadia presented in the manuscript is clear and concise as well. My one big wish for this manuscript is a demonstration of its power in the situations described in the Introduction (lines 69-74), in particular, for those subduction zones where the connection between tremor and slow slip is not straightforward. To that end, I think the biggest test missing from this manuscript is a test in one of the problematic subduction zones, such as New Zealand or Alaska. Much of the demonstration of the power of this technique lies in selecting the threshold. For instance, how applicable are the thresholds determined from northern Cascadia data to other regions (say NZ or AK)? If those subduction zones were examined, how would one go about selecting a threshold? In short, is it possible to select a threshold in the absence of a catalog or does threshold selection depend on slow slip events already being know from other methods?}

\bigskip

We added a section about the analysis of GPS data from New Zealand to show how the method could still be applied if tremor data are not available. \\

To select the threshold in the absence of a tremor catalog, a first possibility is to fit the value of the threshold using data from a subduction zone where tremor data are available. That is what we did in the manuscript: we use for New Zealand the same threshold as we use for northern Cascadia. Another possibility is to try different thresholds, and keep the one that give the most plausible results, with some slow slip events detected but without too much transients that look like noise.

@@

\bigskip

\textit{(Lines 458-460) These lines discuss a small displacement in the time series at 2007.5, but I don't see it. Should I be looking at Figure 9? If so, which latitude band should I look at and which time series are they composed of? That would be the place to look for causal offsets.}

\bigskip

We have added a Figure in the Supplement where we highlight the four possible small displacement that we see on the wavelet details.

@@

\bigskip

\textit{figures 6-8 - I think this figure would be more easily interpretable as well as much effective if only one panel was used, with alternating geodetic and seismic data for a given location (red triangle from Figure 4)in two different colors (black and something else). The red/blue rectangles that indicate detections could be four similar colors for geodetic/seismic detections (blue/purple for negative peaks and red/orange for positive peaks for example). I'd also suggest that figures 7 and 8 could be made into supplementary figures as they show similar information to figure 6}

\bigskip

To better compare between events detected with GPS data and events detected with tremor data, we added figures where both set events are added on top on the wavelet details. We believe the figures are thus more easily readable and interpretable.

### 2022-06-29 paper/paper_bssa.tex
Ready to submit

\begin{itemize}
\item We use a wavelet-based signal processing method to detect transients in GNSS data, such as slow slip events.
\item There is a good correlation between detections of slow slip using GNSS data and using tremor data.
\item The method could be applied in regions where no tremor is detected in conjunction with slow slip events.
\end{itemize}

@@

In many places, tectonic tremor is observed in relation to slow slip and can be used as a proxy to study slow slip events of moderate magnitude where surface deformation is hidden in Global Navigation Satellite System (GNSS) noise. However, in subduction zones where no clear relationship between tremor and slow slip occurrence is observed, these methods cannot be applied, and we need other methods to be able to better detect and quantify slow slip. Wavelets methods such as the Discrete Wavelet Transform (DWT) and the Maximal Overlap Discrete Wavelet Transform (MODWT) are mathematical tools for analyzing time series simultaneously in the time and the frequency domain by observing how weighted differences of a time series vary from one period to the next. In this paper, we use wavelet methods to analyze GNSS time series and seismic recordings of slow slip events in Cascadia. We use detrended GNSS data, apply the MODWT transform and stack the wavelet details over several nearby GNSS stations. As an independent check on the timing of slow slip events, we also compute the cumulative number of tremor in the vicinity of the GNSS stations, detrend this signal, and apply the MODWT transform. In both time series, we can then see simultaneous waveforms whose timing corresponds to the timing of slow slip events. We assume that there is a slow slip event whenever there is a positive peak followed by a negative peak in the wavelet signal. We verify that there is a good correlation between slow slip events detected with only GNSS data, and slow slip events detected with only tremor data for northern Cascadia. The wavelet-based detection method detects well events of magnitude higher than 6 as determined by independent event catalogs (e.g. ~\citep{MIC_2019}).

@@

Slow slip events are a new feature discovered in the last two decades in many subduction zones thanks to recordings of the displacement of Earth's surface by dense Global Navigation Satellite System (GNSS) networks ~\citep{VER_2010,SCH_2010,JIA_2012,WAL_2012}. As with ordinary earthquakes, slow slip events represent slip on a fault, for instance the plate boundary between a tectonic plate subducting under another tectonic plate. However, they take a much longer time (several days to several years) to happen relative to ordinary earthquakes. They have a relatively short recurrence time (months to years) compared to the recurrence time of regular earthquakes (up to several hundreds of years), allowing scientists to observe and study many complete event cycles, which is typically not possible to explore with traditional earthquake catalogs ~\citep{BER_2011}. A slow slip event on the plate boundary is inferred to happen when there is a reversal of the direction of motion at GNSS stations, compared to the secular interseismic motion. Slow slip events have been observed in many places ~\citep{BER_2011,AUD_2016}, such as Cascadia ~\citep{BAR_2020}, Nankai ~\citep{NIS_2013}, Alaska ~\citep{LI_2016}, Costa Rica ~\citep{JIA_2012}, Mexico ~\citep{RAD_2012}, and New Zealand ~\citep{WAL_2020}. \\

In many places, tectonic tremor is also observed in relation to slow slip, but it is more abundant in some places ~\citep{HAL_2018}. Tremor is a long (several seconds to many minutes), low amplitude seismic signal, with emergent onsets, and an absence of clear impulsive phases. Tectonic tremor have been explained as a swarm of small, low-frequency earthquakes (LFEs) ~\citep{SHE_2007_nature}, which are small magnitude earthquakes (M $\sim$ 1) for which frequency content (1-10 Hz) is lower than for ordinary earthquakes (up to 20 Hz). In subduction zones such as Nankai and Cascadia, tectonic tremor observations are spatially and temporally correlated with slow slip observations ~\citep{OBA_2002,ROG_2003}. Due to this correlation, these paired phenomena have been called Episodic Tremor and Slip (ETS). However, this is not always the case. For instance, in northern New Zealand, tremor is more challenging to detect, and seems to be located downdip of the slow slip on the plate boundary ~\citep{TOD_2016}. In Alaska, the tremor zone only partially overlaps the long-term slow slip zone and there does not appear to be any temporal correlation between tremor and slow slip occurrence ~\citep{WEC_2016}. \\

In Cascadia, there are robust signals in both GNSS and tremor ~\citep{HAW_2013}. This is also the case in Nankai ~\citep{HIR_2008}, where tiltmeters are used instead of GNSS. It is thus possible to use tremor as a proxy to observe slow slip events that are not directly observed in the GNSS data. For instance, ~\citet{AGU_2009} studied 23 ETS events in Cascadia with more than 50 hours of tectonic tremor. For all these events, they computed both the GPS-estimated moment release and the cumulative number of hours of tectonic tremor recorded. They observed a linear relationship between moment release and number of hours of tremor for ETS events of moment magnitude 6.3 to 6.8. Based on this linear relationship, it is possible to infer the existence of smaller slow slip events of magnitude 5-6 occurring simultaneously with smaller tremor bursts of duration 1 to 50 hours occurring in between the big ETS events, and for which there is no detectable signal in the GPS data. \\

@@

~\citet{SZE_2008} determined the timing and the amplitude of 34 slow slip events throughout the Cascadia subduction zone between 1997 and 2005 using wavelets. They modeled the GPS time series by the sum of a linear trend, annual and biannual sinusoids representing seasonal effects, Heaviside step functions corresponding to earthquakes and hardware upgrades, and a residual signal. They then applied a Gaussian wavelet transform to the residual time series to get the exact timing of slow slip at each GPS station. The idea is that the wavelet transform allows us to analyze the signal both in the time and the frequency domains. A sharp change in the signal will be localized and seen at all levels of the wavelet decomposition, contrary to what happens with the periodic sinusoids of the Fourier transform. \\

@@

Finally, ~\citet{ALB_2019} used hourly water level records from four tide gauges in the Juan de Fuca Straight and the Puget Sound to determine relative vertical displacements associated with ETS events between 1996 and 2011. Their main idea is that the tidal level measured at a given gauge is the sum of a noise component at multiple timescales (tides, ocean and atmospheric noise) and an uplift signal due to the ETS events. The noise component is assumed to be coherent between all tidal gauges, while the tectonic uplift signal is different provided that the gauges are far enough from each other. By stacking the tidal records after removing tides, the uplift signals cancel each other while the noise signal is amplified. By stacking the details of the DWT decomposition, instead of stacking the raw tidal record, each of the components of the noise at different time scales is retrieved and can then be removed from the raw records to obtain the uplift signal. The authors were then able to clearly see a difference in uplift between the two tidal gauges at Port Angeles and Port Townsend. \\

In our study, we use a similar approach to previous studies with a different reasoning. We only stack signals at nearby GPS stations, assuming that the longitudinal displacement due to the ETS events will then be the same at each of the GPS stations considered. We suppose that some of the noise component is different at each GPS station and will be eliminated by the stacking. Finally, we assume that the noise and the longitudinal displacement due to the ETS events and the secular plate motion have different time scales, so that the wavelet decomposition will act as a bandpass filter to retrieve the displacement signal and highlight the ETS events. We use wavelet methods to analyze GPS and tremor recordings of slow slip events in Cascadia. Our objective is to verify that there is a good correlation between slow slip events detected with only GNSS data, and slow slip events detected with only tremor data. We thus want to demonstrate that the wavelet-based detection method can be applied to detect slow slip events that may currently be obscured using standard methods. \\

@@

We focused our study on northwest Washington State. For the GNSS data, we used the GPS time series provided by the Pacific Northwest Geodetic Array, Central Washington University. These are network solutions in ITRF2008 with phase ambiguities resolved. Solutions are computed with JPL/NASA orbits and satellite clocks. North, East, and Vertical directions are available. However, as the direction of the secular plate motion is close to the East direction, we only used the East direction of the GPS time series for the data analysis, as it has the best signal-to-noise ratio. The wavelet method works best with data with zero mean, and no sharp discontinuities; so we use the cleaned dataset, that is GPS times series with linear trends, steps due to earthquakes or hardware upgrades, and annual and semi-annual sinusoids signals simultaneously estimated and removed following ~\citet{SZE_2004}. For the tremor data, we used the tremor catalog from the Pacific Northwest Seismic Network (PNSN) ~\citep{WEC_2010}. \\

@@

The Discrete Wavelet Transform (DWT) is an orthonormal transform that transforms a time series $X_t \left( t = 0, \cdots , N - 1 \right)$ into a vector of wavelet coefficients $W_i \left( i = 0 , \cdots , N - 1 \right)$. If we denote $J$ the level of the wavelet decomposition, and the number of observations is equal to $N = n * 2^J$, where $n$ is some integer higher or equal to 1, the vector of wavelet coefficients can be decomposed into $J$ wavelet vectors $W_j$ of lengths $\frac{N}{2}$, $\frac{N}{4}$, ... , $\frac{N}{2^J}$, and one scaling vector $V_J$ of length $\frac{N}{2^J}$. Each wavelet vector $W_j$ is associated with changes on time scale $\tau_j = dt 2^{j - 1}$, where $dt$ is the time step of the time series, and corresponds to the filtering of the original time series with a filter with nominal frequency interval $\lbrack \frac{1}{dt 2^{j + 1}} ; \frac{1}{dt 2^j} \rbrack$. The scaling vector $V_J$ is associated with averages in time scale $\lambda_J = dt 2^J$, and corresponds to the filtering of the original time series with a filter with nominal frequency interval $\lbrack 0 ; \frac{1}{dt 2^{j + 1}} \rbrack$. Wavelet vectors can be further decomposed into details and smooths, which are more easily interpretable. We define for $j = 1 , \cdots , J$ the $j$th wavelet detail $D_j$, which is a vector of length $N$, and is associated to time scale $\tau_j = dt 2^{j - 1}$. Similarly, we can define for $j = 1 , \cdots , J$ the $j$th wavelet smooth $S_j$, which is a vector of length $N$, and is associated to scales $\tau_{j + 1} = dt 2^{j + 1}$ and higher. The basic idea is to reapply to $W_j$ the wavelet filter that was used to construct $W_j$ from the initial time series $X$. Together, the details and the smooths define the multiresolution analysis (MRA) of $X$:

@@

@@

The DWT and MODWT methods must be used on a continuous time series, without gaps in the recordings. To deal with the gaps in the GNSS recordings, we simply replace the missing values by interpolation. The value for the first day for which data are missing is equal to the mean of the five days before the gap. The value for the last day for which data are missing is equal to the mean of the five days after the gap. The remaining missing values are computed by doing a linear interpolation of the first and the last values and adding a Gaussian noise component with mean zero and standard deviation equal to the standard deviation of the whole time series. The straight line starts at and ends at . We verify how the wavelet details may be affected by looking at a GPS time series without missing values and compared the wavelet details with and without removing some data points. Station PGC5 recorded continuous 1390 days between 2009 and 2013 without any missing values. We first computed the wavelet details without missing values. Then, we removed ten neighboring values, replaced them using the method described above (linear interpolation plus Gaussian noise), and computed the wavelet details with the replaced values. Figure 2 shows a comparison of the two wavelet details for two different locations of the missing values. We can see that there are visible differences in the time series itself, and in the details at the smallest levels of the wavelet decomposition. However, the differences between the wavelet details with and without missing values get smaller and smaller with increasing levels of details, and are barely visible for the levels that are most relevant (levels 6 and above). We thus conclude that we can easily replace the missing values in the GNSS time series without introducing false detections of slow slip events. \\

We then applied the wavelet filtering to real GPS data. Figure 3 shows the longitudinal displacement for GPS station PGC5, located in southern Vancouver Island, the details of the wavelet decomposition for levels 1 to 8, and the smooth. In the data, we can see a sharp drop in displacement whenever there is a documented slow slip event. For levels 5 to 8, which correspond to time scales 16, 32, 64 and 128 days, we can see in the details a positive peak followed by a negative peak whenever there is a drop in displacement in the data. We thus verify that the wavelet method can detect steps in the time series associated with slow slip events. \\

To increase the signal-to-noise ratio and better detect slow slip events, we stack the signal from several neighboring GPS stations. We choose to focus on GPS stations located close enough to the tremor zone to get a sufficiently high amplitude of the slow slip signal. We choose 16 points along the 40 km depth contour of the plate boundary (model from ~\citet{PRE_2003}) with spacing equal 0.1 degree in latitude (red triangles on Figure 4). Then we took all the GPS stations located in a 50 km radius for a given point, compute the wavelet details for the longitudinal displacement of each station, and stack each detail over the GPS stations. We thus have a stacked detail for each level 1 to 10 of the wavelet decomposition. \\

To assess the success of the wavelet decomposition for detecting slow slip events in GPS time series, we validate the approach by comparing to an independent proxy for ETS events. We took all the tremor epicenters located within a 50 km radius centered on one of the 16 locations marked by red triangles on Figure 3. Then we computed the cumulative number of tremor within this circle. Finally, we removed a linear trend from the cumulative tremor count, and applied the wavelet transform. Figure 5 shows an example of the wavelet decomposition for the third northernmost location on Figure 4 (which is closest to GPS station PGC5). Contrary to what happens for the GPS data, we see a sharp increase in the time series whenever there is a tremor episode, which translates into a negative peak followed by a positive peak in the wavelet details.

\section{Results}

We stacked the 8th level detail of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 locations indicated in Figure 3. The result is shown in the top panel of Figure 6, where each line represents one of the locations along strike. To better highlight the peaks in the wavelet details, we highlighted in red the time intervals where the amplitude of the stacked detail is higher than a threshold, and in blue the  time intervals where the amplitude of the stacked detail is lower than minus the threshold. To compare the GPS signal with the tremor signal, we plotted the 8th level detail of the wavelet decomposition of the tremor count on the bottom panel of Figure 6. We multiplied by -1 the cumulative tremor count for the wavelet decomposition in order to be able to match positive peaks with positive peaks and negative peaks with negative peaks. In the tremor catalog from the PNSN, there are 17 tremor events with more than 150 hours of tremor recorded. The events are summarized in Table 1. The time of the event is the start date plus half the duration of the event. \\

@@

A similar comparison is shown for the wavelet decomposition of the GPS data and the wavelet decomposition of the tremor count data for the 7th level and the 6th level respectively (Figures 7 and 8). The events are harder to see in the 7th level than in the 8th level, both for the GPS data and the tremor count data. The wavelet decomposition is more noisy for the GPS data between 2010 and 2012, but it does not seem that there are more slow slip events visible in the 7th level. \\

For the 6th level detail, we see an additional event in the South in Fall 2009 that is present both in the GPS and the tremor data. It may correspond to the northern extent of a big ETS event occurring in Fall 2009 south of the study area (event 19 in the ~\citet{MIC_2019} catalog). There are three small signals in the GPS data in Winter 2012, Fall 2017, and Winter 2020 that are not present in the tremor data, and may be false detections. To summarize, we assume that true detections are events present in both GPS and tremor time series, and false detections are events present in the GPS but not in the tremor time series. Then, all the 13 events present on the 8th level detail of the wavelet decomposition are true detections and 14 of the 17 events present on the 6th level detail of the wavelet decomposition are true detections. \\

\section{Discussion}

>>>>

\begin{itemize}
\item We use a wavelet-based signal processing method to detect transients in GNSS data, such as slow slip events.
\item In Northern Cascadia, there is good agreement between detections of slow slip using GNSS data and using tremor data.
\item The method can be applied as a starting point to identify slow slip without prior information. New Zealand is shown as on example.
\end{itemize}

@@

In many places, tectonic tremor is observed in relation to slow slip and can be used as a proxy to study slow slip events of moderate magnitude where surface deformation is hidden in Global Navigation Satellite System (GNSS) noise. However, in subduction zones where no clear relationship between tremor and slow slip occurrence is observed, these methods cannot be applied, and we need other methods to be able to better detect and quantify slow slip. Wavelets methods such as the Discrete Wavelet Transform (DWT) and the Maximal Overlap Discrete Wavelet Transform (MODWT) are mathematical tools for analyzing time series simultaneously in the time and the frequency domain by observing how weighted differences of a time series vary from one period to the next. In this paper, we use wavelet methods to analyze GNSS time series and seismic recordings of slow slip events in Cascadia. We use detrended GNSS data, apply the MODWT transform and stack the wavelet details over several nearby GNSS stations. As an independent check on the timing of slow slip events, we also compute the cumulative number of tremor in the vicinity of the GNSS stations, detrend this signal, and apply the MODWT transform. In both time series, we can then see simultaneous waveforms whose timing corresponds to the timing of slow slip events. We assume that there is a slow slip event whenever there is a positive peak followed by a negative peak in the wavelet signal. We verify that there is a good agreement between slow slip events detected with only GNSS data, and slow slip events detected with only tremor data for northern Cascadia. The wavelet-based detection method effectively detects events of magnitude higher than 6 as determined by independent event catalogs (e.g. ~\citep{MIC_2019}). As a demonstration of using the wavelet analysis in a region without significant tremor, we also analyze GNSS data from New Zealand and detect slow slip events that are spatially and temporally close to those detected previously by other studies.

@@

Slow slip events are new phenomena discovered in the last two decades in many subduction zones thanks to recordings of the displacement of Earth's surface by dense Global Navigation Satellite System (GNSS) networks ~\citep{VER_2010,SCH_2010,JIA_2012,WAL_2012}. As with ordinary earthquakes, slow slip events represent slip on a fault, for instance the plate boundary between a tectonic plate subducting under another tectonic plate. However, they take a much longer time (several days to several years) to happen relative to ordinary earthquakes. They have a relatively short recurrence time (months to years) compared to the recurrence time of regular earthquakes (up to several hundreds of years), allowing scientists to observe and study many complete event cycles, which is typically not possible to explore with traditional earthquake catalogs ~\citep{BER_2011}. A slow slip event on the plate boundary is inferred to happen when there is a reversal of the direction of motion at GNSS stations, compared to the secular interseismic motion. Slow slip events have been observed in many places ~\citep{BER_2011,AUD_2016}, such as Cascadia ~\citep{BAR_2020}, Nankai ~\citep{NIS_2013}, Alaska ~\citep{LI_2016}, Costa Rica ~\citep{JIA_2012}, Mexico ~\citep{RAD_2012}, and New Zealand ~\citep{WAL_2020}. \\

In many places, tectonic tremor is also observed in relation to slow slip, but the spatial agreement between tremor and slow slip may vary along the strike of the plate boundary ~\citep{HAL_2018}. Tremor is a long (several seconds to many minutes), low amplitude seismic signal, with emergent onsets, and an absence of clear impulsive phases. Tectonic tremor have been explained as a swarm of small, low-frequency earthquakes (LFEs) ~\citep{SHE_2007_nature}, which are small magnitude earthquakes (M $\sim$ 1) whose frequency content (1-10 Hz) is lower than for ordinary earthquakes (up to 20 Hz). In subduction zones such as Nankai and Cascadia, tectonic tremor observations agree spatially and temporally with slow slip observations ~\citep{ROG_2003,OBA_2004}. Due to this agreement, these paired phenomena have been called Episodic Tremor and Slip (ETS). However, this is not always the case. For instance, in northern New Zealand, tremor is more challenging to detect, and seems to be located downdip of the slow slip on the plate boundary ~\citep{TOD_2016}. In Alaska, the tremor zone only partially overlaps the long-term slow slip zone and there does not appear to be any temporal agreement between tremor and slow slip occurrence ~\citep{WEC_2016}. \\

In Cascadia, there are robust signals in both slow slip and tremor ~\citep{HAW_2013}. This is also the case in Nankai ~\citep{HIR_2008}, where tiltmeters are used instead of GNSS. It is thus possible to use tremor as a proxy to observe slow slip events that are not directly observed in the GNSS data. For instance, ~\citet{AGU_2009} studied 23 ETS events in Cascadia with more than 50 hours of tectonic tremor. For all these events, they computed both the GPS-estimated moment release and the cumulative number of hours of tectonic tremor recorded. They observed a linear relationship between moment release and number of hours of tremor for slow slip events of moment magnitude 6.3 to 6.8. Based on this linear relationship, it is possible to infer the existence of smaller slow slip events of magnitude 5-6 occurring simultaneously with smaller tremor bursts of duration 1 to 50 hours occurring in between the big ETS events, and for which there is no detectable signal in the GPS data. \\

@@

~\citet{SZE_2008} determined the timing and the amplitude of 34 slow slip events throughout the Cascadia subduction zone between 1997 and 2005 using wavelets. They modeled the GPS time series by the sum of a linear trend, annual and biannual sinusoids representing seasonal effects, Heaviside step functions corresponding to earthquakes and hardware upgrades, and a residual signal. They then applied a Gaussian wavelet transform to the residual time series to get the exact timing of slow slip at each GPS station. The idea is that the wavelet transform allows us to analyze the signal both in the time and the frequency domains. A sharp change in the signal will be localized and seen at all time scales of the wavelet decomposition, contrary to what happens with the periodic sinusoids of the Fourier transform. \\

@@

Finally, ~\citet{ALB_2019} used hourly water level records from four tide gauges in the Juan de Fuca Straight and the Puget Sound to determine relative vertical displacements associated with slow slip events between 1996 and 2011. Their main idea is that the tidal level measured at a given gauge is the sum of a noise component at multiple timescales (tides, ocean and atmospheric noise) and an uplift signal due to the slow slip events. The noise component is assumed to be coherent between all tidal gauges, while the tectonic uplift signal is different provided that the gauges are far enough from each other. By stacking the tidal records after removing tides, the uplift signals cancel each other while the noise signal is amplified. By stacking the components at different time scales of the DWT decomposition, instead of stacking the raw tidal record, each of the components of the noise at different time scales is retrieved and can then be removed from the raw records to obtain the uplift signal. Due to the relative location of the tidal gauges at Port Angeles and Port Townsend compared to the slow slip region on the plate boundary, a slow slip event should result in uplift in Port Angeles (western part) and in subsidence in Port Townsend (eastern part). Indeed, the authors were able to clearly see a difference in the sign of the uplift at these two tidal gauges. \\

In our study, we use a similar approach to previous studies with a different reasoning. We only stack signals at nearby GPS stations, assuming that the east-west displacement due to the slow slip events will then be the same at each of the GPS stations considered. We suppose that some of the noise component is different at each GPS station and will be eliminated by the stacking. Finally, we assume that the noise and the longitudinal displacement due to the slow slip events and the secular plate motion have different time scales, so that the wavelet decomposition will act as a bandpass filter to retrieve the displacement signal and highlight the slow slip events. We use wavelet methods to analyze GPS and tremor recordings of slow slip events in Cascadia. Our objective is to verify that there is a good agreement between slow slip events detected with only GNSS data, and slow slip events detected with only tremor data. We thus want to demonstrate that the wavelet-based detection method can be applied to detect slow slip events that may currently be obscured using standard methods. Finally, we apply the method to GNSS data in New Zealand and successfully detect several slow slip events without needing to rely on the tremor data. \\

@@

We first focused our study on northwest Washington State. For the GNSS data, we used the GPS time series provided by the Pacific Northwest Geodetic Array, Central Washington University. These are network solutions in ITRF2014 with phase ambiguities resolved with wide-lane phase-biases. Orbits and satellite clocks provided by the Jet Propulsion Laboratory/NASA. North, East, and Vertical directions are available. However, as the direction of the secular plate motion is close to the East direction, we only used the East direction of the GPS time series for the data analysis, as it has the best signal-to-noise ratio. The wavelet method works best with data with zero mean, and no sharp discontinuities; so we use the cleaned dataset, that is GPS times series with linear trends, steps due to earthquakes or hardware upgrades, and annual and semi-annual sinusoids signals simultaneously estimated and removed following ~\citet{SZE_2004}. For the tremor data, we used the tremor catalog from the Pacific Northwest Seismic Network (PNSN) ~\citep{WEC_2010}. \\

For the application to slow slip events in New Zealand, we used the GPS time series provided by the Geological hazard information for New Zealand (GeoNet). The coordinates have been extracted by GeoNet during the GLOBK run from the combined daily solution files, and converted to (east, north, up) displacement in millimeters with respect to an a priori position and epoch in the ITRF2008 realization. The time series provided by GeoNet have no adjustments made to them, so they may, for example, contain offsets due to earthquakes, offsets due to equipment changes at individual sites, and seasonal (annual and semi-annual) signals due to various causes. Here again, the direction of the secular interseismic plate motion is close to the West direction, so we only used the East-West component of the GPS time series for the data analysis. We detrended the data before applying the wavelet transform by carrying a linear regression of the whole time series and removing the straight line obtained from the regression.

@@

The Discrete Wavelet Transform (DWT) is an orthonormal transform that transforms a time series $X_t \left( t = 0, \cdots , N - 1 \right)$ into a vector of wavelet coefficients $W_i \left( i = 0 , \cdots , N - 1 \right)$. If we denote $J$ the level of the wavelet decomposition, and the number of observations is equal to $N = n * 2^J$, where $n$ is some integer greater than or equal to 1, the vector of wavelet coefficients can be decomposed into $J$ wavelet vectors $W_j$ of lengths $\frac{N}{2}$, $\frac{N}{4}$, ... , $\frac{N}{2^J}$, and one scaling vector $V_J$ of length $\frac{N}{2^J}$. Each wavelet vector $W_j$ is associated with changes on time scale $\tau_j = dt 2^{j - 1}$, where $dt$ is the time step of the time series, and corresponds to the filtering of the original time series with a filter with nominal frequency interval $\lbrack \frac{1}{dt 2^{j + 1}} ; \frac{1}{dt 2^j} \rbrack$. The scaling vector $V_J$ is associated with averages in time scale $\lambda_J = dt 2^J$, and corresponds to the filtering of the original time series with a filter with nominal frequency interval $\lbrack 0 ; \frac{1}{dt 2^{j + 1}} \rbrack$. Wavelet vectors can be further decomposed into details and smooths, which are more easily interpretable. We define for $j = 1 , \cdots , J$ the $j$th wavelet detail $D_j$, which is a vector of length $N$, and is associated to time scale $\tau_j = dt 2^{j - 1}$. Similarly, we can define for $j = 1 , \cdots , J$ the $j$th wavelet smooth $S_j$, which is a vector of length $N$, and is associated to scales $\tau_{j + 1} = dt 2^{j + 1}$ and higher. The basic idea is to reapply to $W_j$ the wavelet filter that was used to construct $W_j$ from the initial time series $X$. Together, the details and the smooths define the multiresolution analysis (MRA) of $X$:

@@

The boundary conditions at the two edges of the time series will affect the wavelet coefficients. For the MODWT, if we denote $L$ the length of the base wavelet filter used for the wavelet decomposition (in our study, we used a Least Asymmetric wavelet filter of length $L = 8$, see ~\citep{PER_2000}, section 4.8, page 107), the length of the wavelet filter at level $j$ used to compute the wavelet detail $D_j$ is:

\begin{equation*}
L_j = \left( 2^j - 1 \right) \left( L - 1 \right) + 1
\end{equation*}

The wavelet coefficients of the detail al level $j$ affected by the boundary conditions at the edges would then be the coefficients with indices $t = 0 , \cdots , L_j - 2$ or $t = N - L_j + 1 , \cdots, N - 1$ (see ~\citep{PER_2000}, section 5.11, page 199). We get $L_j = 442$ for $j = 6$, $L_j = 890$ for $j = 7$ and $L_j = 1786$ for $j = 8$. In practice, the part of the wavelet details affected by the boundary conditions is much shorter than that. We compared the wavelet details computed when using only the data between 2008 and 2012 and the wavelet details computed when using the entire time series from 2000 to 2021 (Figure S1 in the Supplementary Material). Even at level 8 only about 6 months of data on each side are effected by the boundary conditions.

@@

The DWT and MODWT methods must be used on a continuous time series, without gaps in the recordings. To deal with the gaps in the GNSS recordings, we simply replace the missing values by interpolation. The value for the first day for which data are missing is equal to the mean of the five days before the gap. The value for the last day for which data are missing is equal to the mean of the five days after the gap. The remaining missing values are computed by doing a linear interpolation of the first and the last values and adding a Gaussian noise component with mean zero and standard deviation equal to the standard deviation of the whole time series. We verify how the wavelet details may be affected by looking at a GPS time series without missing values and compared the wavelet details with and without removing some data points. Station PGC5 recorded continuous 1390 days between 2009 and 2013 without any missing values. We first computed the wavelet details without missing values. Then, we removed ten neighboring values, replaced them using the method described above (linear interpolation plus Gaussian noise), and computed the wavelet details with the replaced values. Figure S2 in the Supplementary Material shows a comparison of the two wavelet details for two different locations of the missing values. We can see that there are visible differences in the time series itself, and in the details at the smallest levels of the wavelet decomposition. However, the differences between the wavelet details with and without missing values get smaller and smaller with increasing levels of details, and are barely visible for the levels that are most relevant (levels 6 and above). We thus conclude that we can easily replace the missing values in the GNSS time series without introducing false detections of slow slip events. \\

We then applied the wavelet filtering to real GPS data. Figure 2 shows the longitudinal displacement for GPS station PGC5, located in southern Vancouver Island, the details of the wavelet decomposition for levels 1 to 8, and the smooth. In the data, we can see a sharp drop in displacement whenever there is a documented slow slip event. For levels 5 to 8, which correspond to time scales 16, 32, 64 and 128 days, we can see in the details a positive peak followed by a negative peak whenever there is a drop in displacement in the data. We thus verify that the wavelet method can detect steps in the time series associated with slow slip events. \\

To increase the signal-to-noise ratio and better detect slow slip events, we stack the signal from several neighboring GPS stations. We choose to focus on GPS stations located close enough to the tremor zone to get a sufficiently high amplitude of the slow slip signal. We choose 16 points along the 40 km depth contour of the plate boundary (model from ~\citet{PRE_2003}) with spacing equal 0.1 degree in latitude (red triangles on Figure 3). Then we took all the GPS stations located in a 50 km radius for a given point, compute the wavelet details for the longitudinal displacement of each station, and stack each detail over the GPS stations. We thus have a stacked detail for each level 1 to 10 of the wavelet decomposition. \\

To assess the success of the wavelet decomposition for detecting slow slip events in GPS time series, we validate the approach by comparing to an independent proxy for slow slip events. We took all the tremor epicenters located within a 50 km radius centered on one of the 16 locations marked by red triangles on Figure 3. Then we computed the cumulative number of tremor within this circle. Finally, we removed a linear trend from the cumulative tremor count, and applied the wavelet transform. Because of the preprocessing applied to the tremor data before that wavelet transform, the measurement unit associated with the corresponding wavelet details is the fraction of tremor in a day divided by the total number of days. The average value is 1 divided by the total number of days. Figure 4 shows an example of the wavelet decomposition for the third northernmost location on Figure 3 (which is closest to GPS station PGC5). Contrary to what happens for the GPS data, we see a sharp increase in the time series whenever there is a tremor episode, which translates into a negative peak followed by a positive peak in the wavelet details.

\section{Application to data from Cascadia}

We stacked the 8th level detail of the wavelet decomposition of the displacement over all the GPS stations located in a 50 km radius of a given point, for the 16 locations indicated in Figure 3. The result is shown in the top panel of Figure 5, where each line represents one of the locations along strike. To better highlight the peaks in the wavelet details, we highlighted in red the time intervals where the amplitude of the stacked detail is higher than a threshold, and in blue the  time intervals where the amplitude of the stacked detail is lower than minus the threshold. To compare the GPS signal with the tremor signal, we plotted the 8th level detail of the wavelet decomposition of the tremor count on the bottom panel of Figure 5. We multiplied by -1 the cumulative tremor count for the wavelet decomposition in order to be able to match positive peaks with positive peaks and negative peaks with negative peaks. In the tremor catalog from the PNSN, there are 17 tremor events with more than 150 hours of tremor recorded. The events are summarized in Table 1. The time of the event is the start date plus half the duration of the event. \\

@@

A similar comparison is shown for the wavelet decomposition of the GPS data and the wavelet decomposition of the tremor count data for the 7th level and the 6th level respectively (Figures 6 and 7). The events are harder to see in the 7th level than in the 8th level, both for the GPS data and the tremor count data. The wavelet decomposition is more noisy for the GPS data between 2010 and 2012, but it does not seem that there are more slow slip events visible in the 7th level. \\

For the 6th level detail, we see an additional event in the South in Fall 2009 that is present both in the GPS and the tremor data. It may correspond to the northern extent of a big ETS event occurring in Fall 2009 south of the study area (event 19 in the ~\citet{MIC_2019} catalog). There are three small signals in the GPS data in Winter 2012, Fall 2017, and Winter 2020 that are not present in the tremor data, and may be false detections. To summarize, we assume that robust detections are events present in both GPS and tremor time series, and false detections are events present in the GPS but not in the tremor time series. Then, all the 13 events present on the 8th level detail of the wavelet decomposition are robust detections and 14 of the 17 events present on the 6th level detail of the wavelet decomposition are robust detections. \\
